---
title: "SU: Data Cleaning"
author: Brenden Eum
date: February 7, 2023
output: 
  html_document: 
    theme: united
    toc: yes
    toc_float: yes
    toc_depth: 2
    number_sections: yes
    code_folding: hide
    df_print: default
---

# Preamble

Start fresh, set seed, load libraries, set directories.

```{r load, message=FALSE, warning=FALSE}

rm(list = ls())
seed = 1337
library(dplyr)      #pipeline %>%

rawdatadir <- '../data' # raw data
datadir <- '../00 pilot/clean_data' # clean data
figdir <- '../00 pilot/figures' # figure folder
tbldir <- '../00 pilot/tables' # figure folder
tempdir <- '../00 pilot/temp' # temp folder
textdir <- "../00 pilot/text" # text folder
final.figdir <- "../surprise-utility-manuscript/figures"
final.txtdir <- "../surprise-utility-manuscript/text-files"
final.tbldir <- "../surprise-utility-manuscript/tables"

```


# Load Data

* Read the csv file with all the subject data. 
* Make a list of variables that you're interested in, and only keep those in the main data set.
* Only keep subjects that completed the task.

```{r}

variables_of_interest <- c(
  "Participant.Private.ID", #   Unidentifiable subject ID number
  "display", #                  Display name from Gorilla
  "Screen.Name", #              Screen name from Gorilla
  "task", #                     Which task is this trial a part of? (1, 2, 3)
  "block", #                    If it's task 1 (T1), then high or low probability block (H, L).
  "Response", #                 Subject's response
  "Reaction.Time", #            Reaction time in ms
  "Correct", #                  Did subject's response match the correct answer (1=Y, 0=N)
  "correctResponse", #          What was the correct response on this trial? (left or right)
  "pie", #                      What pie graph was shown on the T1 info screen?
  "reward", #                   Was this T1 trial rewarded? (1=Y, 0=N)
  "fractalInfo", #              What fractal was displayed on the T1 info screen?
  "fractalOut", #               What fractal was displayed on the T1 outcome screen?
  "leftInfo", #                 What was displayed on the left side of the T1 info screen?
  "rightInfo", #                What was displayed on the right side of the T1 info screen?
  "leftOut", #                  What was displayed on the left side of the T1 outcome screen?
  "rightOut", #                 What was displayed on the right side of the T1 outcome screen?
  "Response.Type", #            This column & "Response" were to store the mapping: trial condition -> fractal. See bottom of each subject's data.
  "leftFractal", #              What fractal was displayed on the left side of the T2 choice screen?
  "rightFractal", #             What fractal was displayed on the right side of the T2 choice screen?
  "imgSet" #                    What image was shown during the T3 rating screen?
)

data.raw <- read.csv(
  file = file.path(rawdatadir, "pilotdata_exp_106287-v23_task-s39a.csv")
)

data.main <- data.raw[data.raw$Participant.Status=="complete", variables_of_interest]

```


# data.main Cleaning

* Rename columns.
* Factorize.
* Drop NA.

```{r}

# Rename certain columns.

data.main <- data.main %>%
  rename(
    subject = "Participant.Private.ID",
    screen = "Screen.Name",         
    response = "Response",             
    rt = "Reaction.Time",        
    correct = "Correct",            
    correctResponse = "correctResponse",
    T1.pie = "pie",
    T1.reward = "reward",
    T1.fractalInfo = "fractalInfo",
    T1.fractalOut = "fractalOut",
    T1.leftInfo = "leftInfo",
    T1.rightInfo = "rightInfo",
    T1.leftOut = "leftOut",
    T1.rightOut = "rightOut",
    mapping = "Response.Type",
    T2.leftFractal = "leftFractal",
    T2.rightFractal = "rightFractal",
    T3.img = "imgSet"
  )

# Factorize subject number.

data.main$subject <- data.main$subject %>%
  factor() %>%
  as.numeric() %>%
  factor()

```


# data.main Quality Filter

* Drop if...
  * <90% accuracy in T1.
  * Dropped Subject 9 bc their accuracy was 77%.
  
```{r}

# Drop if <70% accuracy in T1.

data.main <- data.main %>%
  group_by(subject, task, display, screen) %>%
  mutate(
    prepre_keep = as.numeric(sum(correct)>=(.9*n())), # We want to calculate keep for T1Trial Outcome screens.
    pre_keep = 0
  )
data.main[data.main$display=="T1Trial" & data.main$screen=="Outcome", "pre_keep"] <- 
  data.main[data.main$display=="T1Trial" & data.main$screen=="Outcome", "prepre_keep"]
data.main <- data.main %>%
  group_by(subject) %>%
  mutate(
    keep = max(pre_keep)
  )
dropped_subjects <- unique(data.main[data.main$keep==0, "subject"])
data.main <- data.main[data.main$keep==1,]
data.main <- select(data.main, -c("prepre_keep", "pre_keep", "keep"))

```


# data.main Fractal Nomenclature

* Examples of fractal nomenclature: fr1p13 means F(R, P=1/3). fr0p0 means F(nR, P=0). fp13H means F(P=1/3, H).
* For T2, make the chosen fractal.

```{r}

# Generate a dataframe that maps from fractal to fractal nomenclature for every subject.

mapping <- data.main[!data.main$mapping %in% c("", "Bonus Trial", "Bonus Correct", "Bonus Response"), ]
mapping <- mapping[ , c("subject","response","mapping")] %>%
  rename(
    fractal = 'response',
    map = "mapping"
  )
mapping[mapping$map=="Hpie1", 'map'] <- "fp1H"
mapping[mapping$map=="Hpie13", 'map'] <- "fp13H"
mapping[mapping$map=="Lpie13", 'map'] <- "fp13L"
mapping[mapping$map=="rewardpie1", 'map'] <- "fr1p1"
mapping[mapping$map=="norewardpie0", 'map'] <- "fr0p0"
mapping[mapping$map=="rewardpie13", 'map'] <- "fr1p13"
mapping[mapping$map=="norewardpie13", 'map'] <- "fr0p13"


# For every row, use the fractal mapping to relabel the T2 left fractal and right fractal.

for (row in c(1:nrow(data.main))) {
  
  currentRow <- data.main[row,]
  
  if (currentRow$display == "T2Trial" & currentRow$screen == "Choice") {
    
    currentSubject <- currentRow$subject
    currentLeftFractal  <- currentRow$T2.leftFractal
    currentRightFractal <- currentRow$T2.rightFractal

    data.main[row, "T2.leftFractal"]  <- mapping[mapping$subject==currentSubject & mapping$fractal==currentLeftFractal, "map"]
    data.main[row, "T2.rightFractal"] <- mapping[mapping$subject==currentSubject & mapping$fractal==currentRightFractal, "map"]
    
  }
}

```


# Save data.main

* Write data.main to a csv.

```{r}

write.csv(
  data.main, 
  file.path(datadir, "data.main.csv")
)

```


# Split data.main Task-wise

```{r}

data.T1 <- data.main[data.main$task==1, ]
data.T2 <- data.main[data.main$task==2, ]
data.T3 <- data.main[data.main$task==3, ]

```


# data.T1 Cleaning

* Drop NA.
* Trial numbers.
* Write data.T1 to a csv.

```{r}

# Only keep relevant columns.

variables_of_interest <- c(
  "subject",
  "task",
  "block",
  "display",
  "screen",
  "rt",
  "response",
  "correct",
  "correctResponse",
  "T1.pie",
  "T1.reward",
  "T1.fractalInfo",
  "T1.fractalOut"
)
data.T1 <- data.T1[,variables_of_interest]

# Drop NA.

data.T1 <- data.T1 %>% na.omit()

# The Outcome screen records all trial related information, so we only need to keep those observations.

display.name <- "T1Trial"
screen.name <- "Outcome"
data.T1 <- data.T1[ data.T1$display == display.name & data.T1$screen == screen.name , ]

# Add a trial number.

data.T1 <- data.T1 %>%
  group_by(subject) %>%
  mutate(
    trial = row_number()
  )

# Write to csv.

write.csv(
  data.T1, 
  file.path(datadir, "data.T1.csv")
)

```


# data.T2 Cleaning

* Write data.T2 to a csv.

```{r}

# Only keep relevant columns.

variables_of_interest <- c(
  "subject",
  "task",
  "display",
  "screen",
  "rt",
  "response",
  "T2.leftFractal",
  "T2.rightFractal"
)
data.T2 <- data.T2[,variables_of_interest]

# Drop NA.

data.T2 <- data.T2 %>% na.omit()

# The Choice screen records all trial related information, so we only need to keep those observations.

display.name <- "T2Trial"
screen.name <- "Choice"
data.T2 <- data.T2[ data.T2$display == display.name & data.T2$screen == screen.name , ]

# Add a trial number.

data.T2 <- data.T2 %>%
  group_by(subject) %>%
  mutate(
    trial = row_number()
  )

# Make a variable for the chosen fractal.

data.T2$T2.chosenFractal <- ""
data.T2[data.T2$response=="left", "T2.chosenFractal"] <- data.T2[data.T2$response=="left", "T2.leftFractal"]
data.T2[data.T2$response=="right", "T2.chosenFractal"] <- data.T2[data.T2$response=="right", "T2.rightFractal"]

# Write to csv.

write.csv(
  data.T2, 
  file.path(datadir, "data.T2.csv")
)

```


# data.T3 Cleaning

* Write data.T3 to a csv.

```{r}

# Only keep relevant columns.

variables_of_interest <- c(
  "subject",
  "task",
  "display",
  "screen",
  "rt",
  "response",
  "T3.img"
)
data.T3 <- data.T3[,variables_of_interest]

# Drop NA.

data.T3 <- data.T3 %>% na.omit()

# The Choice screen records all trial related information, so we only need to keep those observations.

display.name <- "T3Trial"
screen.name <- "Rating Screen"
data.T3 <- data.T3[ data.T3$display == display.name & data.T3$screen == screen.name , ]

# Make response a numerical variable.
data.T3[data.T3$response=="3 (a lot)", "response"] <- toString(3)
data.T3[data.T3$response=="2 (ok)", "response"] <- toString(2)
data.T3[data.T3$response=="1 (little)", "response"] <- toString(1)
data.T3[data.T3$response=="0 (neutral)", "response"] <- toString(0)
data.T3[data.T3$response=="-1 (dislike)", "response"] <- toString(-1)
data.T3$response <- data.T3$response %>% as.integer()

# Add a trial number.

data.T3 <- data.T3 %>%
  group_by(subject) %>%
  mutate(
    trial = row_number()
  )

# Write to csv.

write.csv(
  data.T3, 
  file.path(datadir, "data.T3.csv")
)

```