---
title: "SU: Pilot Visualization"
author: Brenden Eum
date: February 7, 2023
output: 
  html_document: 
    theme: united
    toc: yes
    toc_float: yes
    toc_depth: 2
    number_sections: yes
    code_folding: hide
    df_print: default
---

# Preamble

```{r}

rm(list = ls())
seed = 1337
library(dplyr) #pipeline %>%
library(plotrix) #std.error
library(ggplot2) #plots... duh
library(grid) #combine plots
library(gridExtra) #more combining plots
library(ggsci) #I like the Futurama color palette <3
library(brms) #hierarchical regressions
library(runjags) #MCMC
library(effsize) #cohen's big fat d
library(latex2exp) #for fancy math stuffs in the plots
library(ggpubr) #column titles for grid.arrange plots

datadir <- '../data' #data
figdir <- '../00_pilot/figures' #figures
outdir <- '../00_pilot/output' #output
tempdir <- '../00_pilot/temp' #temp folder

```


# Plot Options

These are ggplot features that you want to stay constant across all your plots. They serve as a default, but don't worry, you can override them anytime if you add the same argument!

```{r plotOptions}

line_width = 1.5
dotsize = .5
jitter_amt = 0.075
textsize <- 18
legtxtsize <- 18
plot_margins <- 1
legend_location <- "none"

col_title_size = 22
col_title_face = 'bold'

back_fill_e = 'lightcyan1'
back_fill_c = 'mistyrose1'
back_fill_j = 'lightcyan2'

plot_width = 8
plot_height = 8
final_plot_units = c('in')

ggplot <- function(...) ggplot2::ggplot(...) + 
  coord_cartesian(expand=F) +
  theme_bw() +
  #plot options
  theme(
    panel.grid.major=element_blank(), 
    panel.grid.minor=element_blank(),
    text=element_text(size=textsize),
    plot.margin=unit(c(plot_margins*.5,plot_margins,plot_margins*.5,plot_margins),"cm")
  ) +
  #legend options
  theme(
    legend.title = element_blank(),
    legend.position = legend_location,
    legend.justification = c(0, 1),
    legend.text = element_text(size = legtxtsize),
    legend.background = element_rect(fill = "transparent")
  ) +
  #color palette
  scale_color_npg() +
  scale_fill_npg()

```


# Read In Data

```{r}

data.T1 <- read.csv(file.path(outdir, "data.T1.csv"))
data.T2 <- read.csv(file.path(outdir, "data.T2.csv"))
data.T3 <- read.csv(file.path(outdir, "data.T3.csv"))

```


# T1: Basic Psychometrics

```{r}

#--------------------------------------------------------------------------------
# Response accuracy.
#--------------------------------------------------------------------------------

pdata.T1.responseAccuracy <- data.T1 %>%
  group_by(subject) %>%
  summarize(
    accuracy = sum(correct)/max(trial)
  )

p.T1.responseAccuracy <- ggplot(pdata.T1.responseAccuracy, aes(accuracy)) +
  # plot data
  annotate(
    "rect", 
    xmin = .86, xmax = .9, 
    ymin = 0, ymax = 50,
    alpha = .1, fill = "darkred") +
  geom_vline(
    xintercept = 0.90, 
    linetype = "dashed", 
    color = "darkred",
    alpha = .7,
    size = line_width
  ) +
  geom_density(
    color = "grey30",
    fill = "grey50",
    size = line_width
  ) +
  # options
  xlim(c(.86,1)) +
  ylim(c(0,50)) +
  labs(
    x = "Accuracy",
    y = "Density"
  )

plot(p.T1.responseAccuracy)

ggsave(
  file.path(figdir, "p.T1.responseAccuracy.pdf"),
  plot = p.T1.responseAccuracy
)

#--------------------------------------------------------------------------------
# Response time.
#--------------------------------------------------------------------------------

pdata.T1.responseTime <- data.T1 %>%
  group_by(subject) %>%
  summarize(
    rt.mean = mean(rt)
  )

p.T1.responseTime <- ggplot(pdata.T1.responseTime, aes(y=rt.mean)) +
  # plot data
  geom_boxplot(
    color = "grey30",
    fill = "grey50",
    size = line_width
  ) +
  # options
  xlim(c(-.9,.9)) +
  ylim(c(0, 1000)) +
  labs(
    y = "Mean Response Time (ms)"
  ) +
  theme(
    axis.title.x=element_blank(),
    axis.text.x=element_blank(),
    axis.ticks.x=element_blank()
  ) +
  coord_cartesian(expand=T)

plot(p.T1.responseTime)

ggsave(
  file.path(figdir, "p.T1.responseTime.pdf"),
  plot = p.T1.responseTime
)

#--------------------------------------------------------------------------------
# Response time by correct.
#--------------------------------------------------------------------------------

pdata.T1.responseTimeByCorrect <- data.T1 %>%
  group_by(subject, correct) %>%
  summarize(
    rt.mean = mean(rt)
  )

pdata.T1.responseTimeByCorrect$correct <- pdata.T1.responseTimeByCorrect$correct %>%
  factor(
    levels = c(1,0),
    labels = c("Correct", "Incorrect")
  )

p.T1.responseTimeByCorrect <- ggplot(pdata.T1.responseTimeByCorrect, aes(x=correct, y=rt.mean, fill = correct)) +
  # plot data
  geom_boxplot(
    color = "grey30",
    size = line_width
  ) +
  # options
  ylim(c(0, 1000)) +
  labs(
    y = "Mean Response Time (ms)"
  ) +
  theme(
    axis.title.x=element_blank()
  ) +
  coord_cartesian(expand=T)

plot(p.T1.responseTimeByCorrect)

ggsave(
  file.path(figdir, "p.T1.responseTimeByCorrect.pdf"),
  plot = p.T1.responseTimeByCorrect
)

```
Response time by correct looks like low time pressure results from Milosavljevic, ..., Koch, and Rangel (2010), where errors are faster than correct responses.


# T2: Quality Check

* Count nonmissing responses by subject.

```{r}

pdata.T2.nonmissingData <- data.T2 %>%
  mutate(
    nonmissing = as.numeric(response != "")
  ) %>%
  group_by(subject) %>%
  summarize(
    obsCount = sum(nonmissing)
  )

p.T2.nonmissingData <- ggplot(pdata.T2.nonmissingData, aes(obsCount)) +
  # plot data
  geom_histogram(
    color = "grey30",
    fill = "grey50",
    size = line_width,
    breaks = c(1:24)
  ) +
  # options
  xlim(c(0,24)) +
  ylim(c(0,8)) +
  labs(
    x = "Task 2 Observation Count (Max=24)",
    y = "Number of Subjects"
  )

plot(p.T2.nonmissingData)

ggsave(
  file.path(figdir, "p.T2.nonmissingData.pdf"),
  plot = p.T2.nonmissingData
)

```


# T2: Evidence of Conditioning

```{r}

# Ideally, we'd like to test F(R, P=1) > F(nR, P=0). I think this is a pretty clear test if people learned.


```


# T2: Model Predictions

* 1:  F(R, P=1/3)  >  F(R, P=1)
* 2:  F(nR, P=0)   >  F(nR, P=1/3)
* 3:  F(P=1, H)    >  F(P=1/3, H)
* 4:  F(P=1/3, L)  >  F(P=1/3, H)
* 5:  F(P=1/3, L)  =  F(nR, P=0)
* 6:  F(P=1/3, H)  =  F(nR, P=1/3)

Examples of fractal nomenclature: fr1p13 means F(R, P=1/3). fr0p0 means F(nR, P=0). fp13H means F(P=1/3, H). 

## * Encode model predictions row-wise

```{r}

# Placeholders
data.T2$pred_index <- NA
data.T2$pred_outcome <- 0

# Test model predictions row-wise.

for (row in c(1:nrow(data.T2))) {
  
  currentRow <- data.T2[row, ]
  
  # 1: F(R, P=1/3) > F(R, P=1)

  fset <- c("fr1p13", "fr1p1")
  if (currentRow$T2.leftFractal %in% fset & currentRow$T2.rightFractal %in% fset) {
    data.T2[row, "pred_index"] <- 1
    if (currentRow$T2.chosenFractal == "fr1p13") {
      data.T2[row, "pred_outcome"] <- 1
    }
  }
  
  # 2: F(nR, P=0) > F(nR, P=1/3)
  
  fset <- c("fr0p0", "fr0p13")
  if (currentRow$T2.leftFractal %in% fset & currentRow$T2.rightFractal %in% fset) {
    data.T2[row, "pred_index"] <- 2
    if (currentRow$T2.chosenFractal == "fr0p0") {
      data.T2[row, "pred_outcome"] <- 1
    }
  }
  
  # 3: F(P=1, H) > F(P=1/3, H)
  
  fset <- c("fp1H", "fp13H")
  if (currentRow$T2.leftFractal %in% fset & currentRow$T2.rightFractal %in% fset) {
    data.T2[row, "pred_index"] <- 3
    if (currentRow$T2.chosenFractal == "fp1H") {
      data.T2[row, "pred_outcome"] <- 1
    }
  }
  
  # 4: F(P=1/3, L) > F(P=1/3, H)
  
  fset <- c("fp13L", "fp13H")
  if (currentRow$T2.leftFractal %in% fset & currentRow$T2.rightFractal %in% fset) {
    data.T2[row, "pred_index"] <- 4
    if (currentRow$T2.chosenFractal == "fp13L") {
      data.T2[row, "pred_outcome"] <- 1
    }
  }
  
  # 5: F(P=1/3, L) = F(nR, P=0)
  
  fset <- c("fp13L", "fr0p0")
  if (currentRow$T2.leftFractal %in% fset & currentRow$T2.rightFractal %in% fset) {
    data.T2[row, "pred_index"] <- 5
    if (currentRow$T2.chosenFractal == "fp13L") {
      data.T2[row, "pred_outcome"] <- 1
    }
  }
  
  # 6: F(P=1/3, H) = F(nR, P=1/3)
  
  fset <- c("fp13H", "fr0p13")
  if (currentRow$T2.leftFractal %in% fset & currentRow$T2.rightFractal %in% fset) {
    data.T2[row, "pred_index"] <- 6
    if (currentRow$T2.chosenFractal == "fp13H") {
      data.T2[row, "pred_outcome"] <- 1
    }
  }
  
}

```

## * Generate plot

```{r}

# 1: F(R, P=1/3) > F(R, P=1)
# 2: F(nR, P=0) > F(nR, P=1/3)
# 3: F(P=1, H) > F(P=1/3, H)
# 4: F(P=1/3, L) > F(P=1/3, H)
# 5: F(P=1/3, L) = F(nR, P=0)
# 6: F(P=1/3, H) = F(nR, P=1/3)

pdata.T2.modelPredictions <- data.T2 %>%
  group_by(subject, pred_index) %>%
  summarize(
    proportion_consistent = sum(pred_outcome)/4
  )

pdata.T2.modelPredictions$pred_index <- factor(pdata.T2.modelPredictions$pred_index)

p.T2.modelPredictions <- ggplot(pdata.T2.modelPredictions, aes(x=pred_index, y=proportion_consistent, fill = pred_index)) +
  # plot data
  geom_hline(
    yintercept = 0.5,
    size = line_width,
    color = "darkred",
    alpha = .2,
    linetype = "dashed"
  ) +
  geom_violin(
    color = "grey30",
    size = line_width
  ) +
  stat_summary(
    fun.data = "mean_se",
    fun.args = list(mult=1),
    geom = "pointrange",
    color = "white"
  ) +
  # options
  labs(
    x = "",
    y = "Proportion of Responses"
  ) +
  coord_cartesian(expand=T) +
  scale_y_continuous(
    breaks = c(0, .5, 1)
  ) +
  scale_x_discrete(
    breaks = c(1:6),
    labels = c(
      bquote(bold("F(R, P=1/3)")~"> F(R, P=1)"),
      bquote(bold("F(nR, P=0)")~"> F(nR, P=1/3)"),
      bquote(bold("F(P=1, H)")~"> F(P=1/3, H)"),
      bquote(bold("F(P=1/3, L)")~"> F(P=1/3, H)"),
      bquote(bold("F(P=1/3, L)")~"= F(nR, P=0)"),
      bquote(bold("F(P=1/3, H)")~"= F(nR, P=1/3)")
    )
  ) +
  theme(
    axis.text.x = element_text(
      size = 10.5,
      angle = 15, 
      vjust = 1, 
      hjust = 1
    )
  )

plot(p.T2.modelPredictions)

ggsave(
  file.path(figdir, "p.T2.modelPredictions.pdf"),
  plot = p.T2.modelPredictions
)

```



# T3: Liking Ratings

```{r}

#--------------------------------------------------------------------------------
# Liking Ratings
#--------------------------------------------------------------------------------

pdata.T3.ratings <- data.T3

p.T3.ratings <- ggplot(pdata.T3.ratings, aes(response)) +
  # plot data
  geom_histogram(
    color = "grey30",
    fill = "grey50",
    size = line_width,
    binwidth = 1
  ) +
  # options
  ylim(c(0,220)) +
  labs(
    x = "Liking Ratings",
    y = "Count"
  ) +
  coord_flip(expand=T)

plot(p.T3.ratings)

ggsave(
  file.path(figdir, "p.T3.ratings.pdf"),
  plot = p.T3.ratings
)

#--------------------------------------------------------------------------------
# Liking Ratings By Subject
#--------------------------------------------------------------------------------

pdata.T3.ratingsBySubject <- data.T3

pdata.T3.ratingsBySubject$subject <- pdata.T3.ratingsBySubject$subject %>%
  factor()

p.T3.ratingsBySubject <- ggplot(pdata.T3.ratingsBySubject, aes(x=subject, y=response, fill = subject)) +
  # plot data
  geom_violin(
    color = "grey30",
    size = line_width
  ) +
  # options
  labs(
    x = "Subject",
    y = "Liking Ratings"
  ) +
  coord_cartesian(expand=T)

plot(p.T3.ratingsBySubject)

ggsave(
  file.path(figdir, "p.T3.ratingsBySubject.pdf"),
  plot = p.T3.ratingsBySubject,
  width = plot_width,
  height = plot_height *.33
)

```